---
title: "Blackbody Radiation: spectral density of electromagnetic radiation emitted by a blackbody"
author: "dB"
date: "8/13/2022"
output: ioslides_presentation
widescreen: true
smaller: true
runtime: shiny
---

```{r setup,  include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Planck's Law
Planck's law examines the spectral density of electromagnetic radiation emitted by a blackbody in thermal equilibrium, at emissivity=1, and given temperature T, where there is zero net flow of matter or energy between the body and its environment.
Planck's Law Equation (Wavelength):
$${\displaystyle B_{\lambda }(\lambda ,T)={\frac {2hc^{2}}{\lambda ^{5}}}{\frac {1}{e^{(hc/\lambda kT)}-1}}}$$
Planck's Law Equation (Frequency):
$${\displaystyle B(\nu ,T)={\frac {2h\nu ^{3}}{c^{2}}}{\frac {1}{e^{{(h\nu/ }{kT})}-1}}}$$

### Wien's Law descibes peak wavelength and peak frequency.

```{r, warning=FALSE, message=FALSE, error=FALSE}
library(stats)
# contants
h <- 6.625 * 10^(-34)
k <- 1.380649 * 10^(-23)
c <- 2.99792458 * 10^8
b_λ  <- 2.897771955 *10^(-3)

# Temperature range
T_7k <- 7000
T_6k <- 6000
T_5k <- 5000
T_4k <- 4000
T_3k <- 3000

# wavelength range in meters
wav_len <- 10^seq(-8, -5.7, 0.01)

# Spectral Density function
B <- function(x, y) {
# spectral density function of B with x= wave length and y= temperature
result <- (2 * h * (c^2)) / (x^(5) * ((exp((h*c) / (x * k * y))) -1))
}

# spectral densities
B_7k <- B(x=wav_len, y=T_7k)
B_6k <- B(x=wav_len, y=T_6k)
B_5k <- B(x=wav_len, y=T_5k)
B_4k <- B(x=wav_len, y=T_4k)
B_3k <- B(x=wav_len, y=T_3k)

# Wien's Law for peak wavelength(maximum) at specific temperature
lambda_7k <- b_λ/T_7k
lambda_6k <- b_λ/T_6k
lambda_5k <- b_λ/T_5k
lambda_4k <- b_λ/T_4k
lambda_3k <- b_λ/T_3k

Wien_7k <- B(x=lambda_7k, y=T_7k)
Wien_6k <- B(x=lambda_6k, y=T_6k)
Wien_5k <- B(x=lambda_5k, y=T_5k)
Wien_4k <- B(x=lambda_4k, y=T_4k)
Wien_3k <- B(x=lambda_3k, y=T_3k)

wav_len_max <- c(lambda_7k, lambda_6k, lambda_5k, lambda_4k, lambda_3k)
#wav_len_max*10^9

Wien_max <- c(Wien_7k, Wien_6k, Wien_5k, Wien_4k, Wien_3k)
#Wien_max
```
## Planck's Law for wavelength with Wien's Law showing peak wavelengths at temperature T(K)
```{r}
plot(wav_len*10^9, B_7k*10^-13, type="l", col="blue", ylab="B, Spectral Radiance, 10^13 * W/m^2/sr/nm", xlab="Wavelength (nm)")
lines(wav_len*10^9, B_6k*10^-13, type="l", col="green")
lines(wav_len*10^9, B_5k*10^-13, type="l", col="yellow")
lines(wav_len*10^9, B_4k*10^-13, type="l", col="orange")
lines(wav_len*10^9, B_3k*10^-13, type="l", col="red")
lines(wav_len_max*10^9, Wien_max*10^-13, type="l", col="purple")
legend('topright', legend=c('7000K','6000K','5000K','4000K', '3000K'), col=c('blue', 'green', 'yellow', 'orange','red'), lty=1)
points(wav_len_max*10^9, Wien_max*10^-13) 
title("Plank's Law and Wien's Law vs Wavelength")
```


```{r, warning=FALSE, message=FALSE, error=FALSE}
library(stats)
# contants
b_ν <- 5.88 * 10^(10)

# Temperature range
T_7k <- 7000
T_6k <- 6000
T_5k <- 5000
T_4k <- 4000
T_3k <- 3000

# wavelength range in THz (10^12 Hz)
freq_span <- 10^seq(9, 15.3, 0.001)

# Spectral Density function vs Frequency
B_freq <- function(x, y) {
# spectral density function of B with x=frequency and y= temperature
#result <- (2 * h * (c^2)) / (x^(5) * ((exp((h*c) / (x * k * y))) -1))
result <- (2 * h * (x^3)) / (c^(2) * ((exp((h*x) / (k * y))) -1))
}

# Spectral Density function span vs Frequency at specific Temperature
B_7k <- B_freq(x=freq_span, y=T_7k)
B_6k <- B_freq(x=freq_span, y=T_6k)
B_5k <- B_freq(x=freq_span, y=T_5k)
B_4k <- B_freq(x=freq_span, y=T_4k)
B_3k <- B_freq(x=freq_span, y=T_3k)

# Wien's Law for peak wavelength(maximum) at specific temperature
freq_7k <- b_ν*T_7k
freq_6k <- b_ν*T_6k
freq_5k <- b_ν*T_5k
freq_4k <- b_ν*T_4k
freq_3k <- b_ν*T_3k
max_B_freq <- c(freq_3k, freq_4k, freq_5k, freq_6k, freq_7k)
#max_B_freq*10^-12

Wien_7k <- B_freq(x=freq_7k, y=T_7k)
Wien_6k <- B_freq(x=freq_6k, y=T_6k)
Wien_5k <- B_freq(x=freq_5k, y=T_5k)
Wien_4k <- B_freq(x=freq_4k, y=T_4k)
Wien_3k <- B_freq(x=freq_3k, y=T_3k)
Wien_max <- c(Wien_3k, Wien_4k, Wien_5k, Wien_6k, Wien_7k)
#Wien_max * 10^8
```

## Planck's Law for frequency with Wien's Law showing peak frequencies at temperature T(K)
```{r}
plot(freq_span*10^-12, B_7k*10^8, type="l", col="blue", ylab="B, Spectral Radiance, 10^8 * W/m^2/sr/THz", xlab="Frequency (THz)")
lines(freq_span*10^-12, B_6k*10^8, type="l", col="green")
lines(freq_span*10^-12, B_5k*10^8, type="l", col="yellow")
lines(freq_span*10^-12, B_4k*10^8, type="l", col="orange")
lines(freq_span*10^-12, B_3k*10^8, type="l", col="red")
lines(max_B_freq*10^-12, Wien_max*10^8, type="l", col="purple")
points(max_B_freq*10^-12, Wien_max*10^8, col = "dark red") 
legend('topright', legend=c('7000K','6000K','5000K','4000K', '3000K'), col=c('blue', 'green', 'yellow', 'orange','red'), lty=1)
title("Plank's Law and Wien's Law vs Frequency")
```

## Blackbody Spectral Radiance Interactive App
```{r, echo = FALSE}
library(shiny)
library(stats)
# contants
h <- 6.625 * 10^(-34)
k <- 1.380649 * 10^(-23)
c <- 2.99792458 * 10^8
b_λ  <- 2.897771955 *10^(-3)
b_ν <- 5.88 * 10^(10)

# wavelength range in meters
wav_len <- 10^seq(-8, -5.5, 0.01)

# wavelength range in THz (10^12 Hz)
freq_span <- 10^seq(9, 15.3, 0.01)

# Spectral Radiance function
B_λ <- function(x, y) {
    # spectral density function of B with x= wave length and y= temperature
    result <- (2 * h * (c^2)) / (x^(5) * ((exp((h*c) / (x * k * y))) -1))
}
B_v <- function(x, y) {
    # spectral density function of B with x=frequency and y= temperature
    #result <- (2 * h * (c^2)) / (x^(5) * ((exp((h*c) / (x * k * y))) -1))
    result <- (2 * h * (x^3)) / (c^(2) * ((exp((h*x) / (k * y))) -1))
}

ui <- fluidPage(
  headerPanel(h3('Blackbody Spectral Radiance')),
  sidebarPanel(
      selectInput(inputId = "Xaxis", 
                  label = "Choose your X-axis Unit", 
                  choices = c("Wavelength (nm)", "Frequency (THz)")),
  
     sliderInput(inputId = "slider1", 
                 label = "Choose a Temperature (Kelvin)", 
                 value = 5000, min = 0, max = 10000)),

            mainPanel(
                h4("Temperature (Kelvin)"),
                h3(textOutput("text0")),
                h4(textOutput("text1")),
                h3(textOutput("text2")),
                h4("Peak Wavelength (nm)"),
                h3(textOutput("text3")),
               h4("Peak Frequency (THz)"),
                h3(textOutput("text4")),
            plotOutput('plot1')
  ))


server <- function(input, output) {
        

            output$text0 <- renderText(input$slider1)
        
        output$text1 <- renderText({
            
            if (input$Xaxis == "Wavelength (nm)") {
                "Peak Spectral Radiance (W/m^2/sr/nm)"}
            
            else if (input$Xaxis == "Frequency (THz)"){ 
                "Peak Spectral Radiance (W/m^2/sr/THz)"}
        })
        
        output$text2 <- renderText({
            
            if (input$Xaxis == "Wavelength (nm)") {
                (B_λ(b_λ/input$slider1, input$slider1))}
            
            else if (input$Xaxis == "Frequency (THz)"){ 
                (B_v(b_ν * input$slider1, input$slider1))*10^(16)}
        })
        
        output$text3 <- renderText((b_λ/input$slider1 * 10^9))
        output$text4 <- renderText((b_ν * input$slider1) * 10^(-12))
      
# plot       
        output$plot1 <- renderPlot({
            
            if (input$Xaxis == "Wavelength (nm)") {
                plot(wav_len*10^9, ((B_λ(wav_len, input$slider1))*10^-13), 
                     type="l", 
                     col="blue", 
                     ylab="B, Spectral Radiance, 10^13 * W/m^2/sr/nm", 
                     xlab="Wavelength (nm)")
                    title("Plank's Law and Wien's Law vs Wavelength")}
            
            else if (input$Xaxis == "Frequency (THz)"){ 
                plot(freq_span*10^-12, ((B_v(freq_span, input$slider1))*10^8), 
                     type="l", 
                     col="blue", 
                     ylab="B, Spectral Radiance, 10^8 * W/m^2/sr/THz", 
                     xlab="Frequency (THz)")
                    title("Plank's Law and Wien's Law vs Frequency")}}
            )}

shinyApp(ui = ui, server = server)
```


